Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-28 01-clean.Rmd
Add gwp cleaning

```{r packages}
#| results: 'hide'
# Packages
library(readxl)
library(imputeTS)
library(naniar)
library(scales)
library(janitor)
library(countrycode)
library(kableExtra)
library(tidyverse)

@@

@@

```{r gbd-harmonise-countries}
# Check what names were changed and if any don't have harmonised counterpart
gbd %>%
  distinct(country) %>% 
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>% 
  filter(country != country_harmonised | is.na(country_harmonised)) %>% 
  kbl() %>% 
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "400px")
# Harmonise old names after verifying above that none are dropped
gbd <- gbd %>% 
  mutate(country = countryname(country, destination = "cldr.short.en"))
```

GWP cleaning goes here.

@@

# Harmonise old names after verifying above that none are dropped
itu <- itu %>% 
  mutate(country = countryname(country, destination = "cldr.short.en"))

# These countries don't exist in outcomes
itu %>%
  distinct(country) %>% 
  anti_join(distinct(gbd, country)) %>% 
  kbl() %>% 
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "400px")

# Drop them
itu <- itu %>% 
  filter(country %in% unique(gbd$country))
```

@@

Remove countries with no observations

```{r}
itu <- itu %>% 
  group_by(country) %>% 
  mutate(nobs = sum(!is.na(internet))) %>% 
  ungroup() %>% 
  filter(nobs > 0) %>% 
  select(-nobs)
```

@@

# Countries in our dataset not present in region data
filter(
  distinct(gbd, country), 
  !(country %in% unique(regions$country))
) %>% 
  arrange(country) %>% 
  write_csv("data-raw/Regions/GBD-countries-without-region.csv")

@@

# Load our manually filled region data
regions_filled <- read_csv(
  "data-raw/Regions/GBD-countries-without-region-filled.csv"
) 
```

@@

```{r}
# Add regions to data tables
itu <- left_join(itu, regions)
gbd <- left_join(gbd, regions)
```

@@

```{r}
gbd_all <- gbd
gbd <- gbd %>% 
  filter(age %in% c("10 to 14", "15 to 19", "20 to 24"))
```

@@

```{r}
write_rds(gbd, "data/gbd.rds")
write_rds(gbd_all, "data/gbd_all_ages.rds")
write_rds(itu, "data/itu.rds")
```

>>>>

```{r packages}
#| results: 'hide'
# Packages
library(readxl)
library(imputeTS)
library(naniar)
library(scales)
library(janitor)
library(haven)
library(labelled)
library(countrycode)
library(kableExtra)
library(tidyverse)

@@

```{r}
# First liberate the data from the slow SPSS file and save to disk
gwp_path <- "data-raw/Gallup/gwp-processed.rds"
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    "data-raw/Gallup/The_Gallup_101521.sav", 
    col_select = c(
      YEAR_CALENDAR, 
      COUNTRYNEW, 
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74
    )
  )

  # Get rid of SPSS attributes
  gwp <- gwp %>% 
    zap_labels() %>% 
    zap_label() %>% 
    zap_widths() %>% 
    zap_formats()

  # Save into a good format
  write_rds(gwp, gwp_path)
} else {gwp <- read_rds(gwp_path)}

# Codes NOTE 1: yes 2: no!
read_spss(
  "data-raw/Gallup/The_Gallup_101521.sav", 
  n_max = 10,
  col_select = c(
    WPID, WP1219,
    WP16,
    WP60, WP61, WP63, WP65, WP67,
    WP68, WP69, WP70, WP71, WP74
  )
) %>% 
  generate_dictionary()

gwp <- gwp %>% 
  clean_names() %>% 
  transmute(
    country = countrynew,
    year = year_calendar,
    id = wpid,
    sex = factor(wp1219, levels = c(2, 1), labels = c("Female", "Male")),
    age = wp1220,
    # Don't know, refused, and missing values
    `Life\nsatisfaction` = if_else(between(wp16, 0, 10), wp16, NaN),
    across(wp60:wp74, ~if_else(between(., 1, 2), ., NaN)),
    # Also reverse the weird 1: yes 2: no coding here
    across(wp60:wp74, ~3 - .)
  )

# Scale scores note rescaling
gwp <- gwp %>% 
  mutate(
    `Life\nsatisfaction` = `Life\nsatisfaction` / 10,
    `Negative\nexperiences` = 
      rowMeans(select(., wp68:wp74), na.rm = TRUE) - 1,
    `Positive\nexperiences` = 
      rowMeans(select(., wp60:wp67), na.rm = TRUE) - 1
  ) %>% 
  select(-c(wp60:wp74))

# Categorize ages
gwp <- gwp %>% 
  mutate(
    age = cut(
      age, 
      breaks = seq(15, 105, by = 5), 
      include.lowest = TRUE,
      right = FALSE,
      labels = paste(seq(15, 100, by = 5), "to", seq(19, 104, by = 5))
    ) %>% as.character()
  )

# There is one 13 year old which resulted in a NA and we drop it
gwp <- drop_na(gwp, age)
```

Summarise gwp data

```{r}
gwp <- gwp %>% 
  pivot_longer(
    c(
      `Life\nsatisfaction`, 
      `Negative\nexperiences`, 
      `Positive\nexperiences`
    ), 
    names_to = "cause", values_to = "val"
  ) %>% 
  group_by(country, year, sex, age, cause) %>% 
  summarise(
    se = sd(val, na.rm = TRUE) / sqrt(n()),
    val = mean(val, na.rm = TRUE)
  ) %>% 
  ungroup()
```

Harmonise countries

```{r gwp-harmonise-countries}
# Check what names were changed and if any don't have harmonised counterpart
gwp %>%
  distinct(country) %>% 
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>% 
  filter(country != country_harmonised | is.na(country_harmonised)) %>% 
  kbl() %>% 
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "400px")
# Harmonise old names and replace only if harmonised name found
gwp <- gwp %>% 
  mutate(
    country_harmonised = countryname(
      country, destination = "cldr.short.en"
    )
  ) %>% 
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>% 
  select(-country_harmonised)

distinct(gwp, country)
distinct(gwp, country) %>% filter(is.na(country))
```

@@

```{r gbd-harmonise-countries}
# Check what names were changed and if any don't have harmonised counterpart
gbd %>%
  distinct(country) %>% 
  arrange(country) %>%
  mutate(
    country_harmonised = countryname(country, destination = "cldr.short.en")
  ) %>% 
  filter(country != country_harmonised | is.na(country_harmonised)) %>% 
  kbl() %>% 
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "400px")
# Harmonise old names and replace only if harmonised name found
gbd <- gbd %>% 
  mutate(
    country_harmonised = countryname(
      country, destination = "cldr.short.en"
    )
  ) %>% 
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>% 
  select(-country_harmonised)

distinct(gbd, country)
distinct(gbd, country) %>% filter(is.na(country))
```

@@

# Harmonise old names and replace only if harmonised name found
itu <- itu %>% 
  mutate(
    country_harmonised = countryname(
      country, destination = "cldr.short.en"
    )
  ) %>% 
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>% 
  select(-country_harmonised)

distinct(itu, country)
distinct(itu, country) %>% filter(is.na(country))

# These countries don't exist in outcomes
itu %>%
  distinct(country) %>% 
  mutate(gwp = country %in% unique(gwp$country)) %>% 
  mutate(gbd = country %in% unique(gbd$country)) %>% 
  filter(!gwp | !gbd) %>% 
  kbl() %>% 
  kable_minimal(full_width = FALSE) %>% 
  scroll_box(width = "500px", height = "400px")

# Drop them
itu <- itu %>% 
  filter(country %in% unique(gbd$country) | country %in% unique(gwp$country))
```

@@

@@

# Countries in GBD not present in region data
no_region_gbd <- filter(
  distinct(gbd, country), 
  !(country %in% unique(regions$country))
) %>% 
  arrange(country)

# Then see what countries in GWP are without region
no_region_gwp <- filter(
  distinct(gwp, country), 
  !(country %in% unique(regions$country))
) %>% 
  arrange(country)

bind_rows(no_region_gbd, no_region_gwp) %>% 
  arrange(country) %>% 
  distinct() %>%
  write_csv("data-raw/Regions/countries-without-region.csv")

@@

# Load our manually filled region data
regions_filled <- read_csv(
  "data-raw/Regions/countries-without-region-filled.csv"
)
```

@@

```{r}
# Add regions to data tables
itu <- left_join(itu, regions)
gbd <- left_join(gbd, regions)
gwp <- left_join(gwp, regions)
```

## Create tables for young people

@@

```{r}
gwp_all <- gwp
gwp <- gwp %>% 
  filter(age %in% c("15 to 19", "20 to 24"))

gbd_all <- gbd
gbd <- gbd %>% 
  filter(age %in% c("10 to 14", "15 to 19", "20 to 24"))
```

@@

```{r}
write_rds(gwp, "data/gwp.rds")
write_rds(gwp_all, "data/gwp_all_ages.rds")
write_rds(gbd, "data/gbd.rds")
write_rds(gbd_all, "data/gbd_all_ages.rds")
write_rds(itu, "data/itu.rds")
```
